Automating Verification of MPC Security

### 2023-06-18 popl24/composition.tex
checkpoint

Our aim is to demonstrated that $\gNIMO$ is preserved by gates within circuits,
where preservation is expressed via pre-~and post-conditions. We can use
techniques in Section \ref{section-bruteforce} to automatically verify postconditions
for components, using witness programs that satisfy preconditions prior to
component execution. For example, we can automatically verify that $\ttt{encode}$
generates a program and wire values the enjoy $\gNIMO$. Then, to verify
that we can verify
that the $\minifed$ program and wire value generated by the following program satisfies
$\gNIMO$:
\begin{verbatimtab}
  let secrets = encode(s1,s2) in
  andgate(0,s1,s2); evalgate(secrets.wv1,secrets.wv2)
\end{verbatimtab}
However, we need to establish that $\ttt{encode(s1,s2)}$ is an
adequate witness to correctness, so our automated verification generalizes to
arbitrary circuit contexts.

@@

We aim to show that gates are probabilistically independent of
each other (i.e., are separated), and this independence supports
correctness-preserving composition. Noting that gates have
naming conventions, and that gates are distinguished by gate
identifiers $g$, we can define what we mean by a gate
$g$ being in a given program $\prog$, written $g \in \prog$. 
\begin{definition}
  We write $g \in \prog$ iff $\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
  and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
  or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
  and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
\end{definition}

A critical property is that separation is preserved by safe
(in the sense of Theorem \ref{theorem-metalang-safety}) substitution
of programs, assuming disjointness of flips used in programs separated
programs. 
\begin{lemma}
  \label{lemma-scope}
  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and
  $\flips(\prog_1) \cap \flips(\prog_2) = \varnothing$ then
  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$ for any
  $\prog_1'$ where $\prog_1';\prog_2$ is safe and $\flips(\prog_1')
  \cap \flips(\prog_2) = \varnothing$.
\end{lemma}
\begin{proof}
  \cnote{The Lemma statement may be slightly off, there may be more restrictions on $\prog_1$.
  But generally this result follows by results from \cite{barthe2019probabilistic.}}
\end{proof}

Another important property is that $\gNIMO$ is preserved by substitution
for a witness program. This follows by Lemma \ref{lemma-scope} and definition
of $\gNIMO$. 
\begin{lemma}
  \label{lemma-gnimo}
  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
  and $\flips(\prog_1) \cap \flips(\prog_2) = \varnothing$ and
  $\gNIMO(\prog_1;\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ for
  $g \in \prog_1$,
  then $\gNIMO(\prog_1';\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
  for any $\prog_1'$ where $\prog_1';\prog_2$ is safe and
  $\flips(\prog_1') \cap \flips(\prog_2) = \varnothing$.
\end{lemma}

@@

Now we can demonstrate that our $\ttt{and}$ gate definition preserves
$\gNIMO$, and additionally that internale gates are probabilistically independent.
This latter fact is crucial to compositionality, and allows us for automatically
verify the circuit in isolation, with input encoding serving as an adequate input
witness for generalization. 
\begin{lemma}[Composition ($\eand$)]
  \label{lemma-compose}
Given well-typed $\prog_1;E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]$ where:
$$
\config{\prog_1}{E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
$$
If the following preconditions hold:
\begin{enumerate}
\item $g_1 \ne g_2$
\item $\gNIMO(\prog_1,v_1,g_1) \wedge \gNIMO(\prog_1,v_2,g_2)$ 
\end{enumerate}
then we have as postconditions:
\begin{enumerate}
  \item $\gNIMO(\prog_1;\prog_2,v,g)$
  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
\end{enumerate}
\end{lemma}
\begin{proof}
  We have automatically verified the postconditions (1) and (2) for the particular
  program $\prog_1;\prog_2$ resulting from execution of:
  $$e\quad \defeq\quad \ttt{let gs = encode(s1,s2) in andgg(0,s1,s2,gs.wv1,gs.wv2)}$$
  where:
  $$
  \config{\varnothing}{e} \redxs
  \config{\prog_1}{\ttt{andgg}(\ttt{0},\ttt{s1},\ttt{s2},v_1,v_2)} \redxs \config{\prog_1;\prog_2}{v}
  $$
  By Lemma \ref{lemma-encode} we have that
  $\gNIMO(\prog_1,v_1,\ttt{s1}) \wedge \gNIMO(\prog_1,v_2,\ttt{s2})$. 
  And since $\ttt{andgate}$ generates unique flips per gate, and well-typedness ensures
  uniqueness of gates, we can generalize the result to replace $\prog_1$ with arbitrary $\prog_1'$
  such that $\prog_1';\prog_2$ is safe,
  applying Lemma \ref{lemma-scope} to obtain postcondition (2) and Lemma \ref{lemma-gnimo}
  to obtain postcondition (1). 
\end{proof}
Finally, we can demonstrate that decoding correctly casts a garbled output into the
clear, and thus establishes $\NIMO$ for the garbled circuit that precedes it.
\begin{lemma}[Decoding]
  \label{lemma-decode}
Given well-typed $\prog_1;E[\ttt{decode}(g,v)]$ where
$$
\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
$$
If $\gNIMO(\prog_1,v,g)$, then $\NIMO(\prog_1;\prog_2;\eassign{\outv}{\be})$.
\end{lemma}
\begin{proof}
  We have automatically verified $
  \progd(\prog,\be)(\{ \itv \mapsto 1 \}) = 1$ where:
  $$\config{\varnothing}{\ttt{decode(g,\{ k = flip[2, gate:g.k]; p = flip[2,gate:g.p] \})}}
    \redxs \config{\prog}{\be}$$
    and also $
  \progd(\prog,\be)(\{ \itv \mapsto 0 \}) = 1$ where:
  $$\config{\varnothing}{\ttt{decode(g,\{ k = not flip[2, gate:g.k]; p = not flip[2,gate:g.p] \})}}
    \redxs \config{\prog}{\be}$$
  The result follows by definition of $\gNIMO$ and $\NIMO$.
\end{proof}

>>>>

Our aim is to demonstrated that $\gNIMO$ is preserved by gates within circuits,
where preservation is expressed via pre-~and post-conditions. We can use
techniques in Section \ref{section-bruteforce} to automatically verify postconditions
for components, using witness programs that satisfy preconditions prior to
component execution. For example, we can automatically verify that $\ttt{encode}$
generates a program and wire values the enjoy $\gNIMO$. Then, to verify
correctness of $\ttt{and}$ gate construction and evaluation- i.e., that it
preserves $\gNIMO$- we can verify
that the $\minifed$ program and wire value generated by the following program
$\metaprot$ program satisfies $\gNIMO$:
\begin{verbatimtab}
  let secrets = encode(s1,s2) in
  andgate(0,s1,s2); evalgate(secrets.wv1,secrets.wv2)
\end{verbatimtab}
However, we need to establish that $\ttt{encode(s1,s2)}$ is an
adequate witness, in that it ensures that automated verification generalizes to
arbitrary circuit contexts.

@@

\begin{definition}
  We say that $\prog$ \emph{covers} $X$ iff
  $\progd(\prog)(\store) > 0$ for all $\store\in\mems(X)$.
\end{definition}

We aim to show that gates are probabilistically independent of each
other (i.e., are separated), and this independence supports
correctness-preserving composition. A critical property is that
separation is preserved by safe (in the sense of Theorem
\ref{theorem-metalang-safety}) substitution of programs, assuming
disjointness of flips used in programs separated programs.
\begin{lemma}
  \label{lemma-scope}
  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$ and $\prog_1$
  covers $X$, then 
  $(\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_1)}) *
  (\margd{\progd(\prog_1';\prog_2)}{\iov(\prog_2)})$ for any
  $\prog_1'$ such that $\vars(\prog_1') \cap \vars(\prog_2) = X$ and
  $\prog_1';\prog_2$ is safe.
\end{lemma}
\begin{proof}
  Suppose on the contrary that there exists $\prog_1'$ that is not
  separable from $\prog_2$. By results in
  \cite{barthe2019probabilistic}, this implies that there
  exists $\store^n$ and $\store^n$ for $0 \le n \le 1$ such that
  $\store^1_X = \store^{1'}_X$ where:
  $$
  \config{\store^0}{\prog_1;\prog_2} \redxs \config{\store^1}{\prog_2} \redxs
  \config{\store^2}{\varnothing}
  $$
  and:
  $$
  \config{\store^{0'}}{\prog_1;\prog_2} \redxs \config{\store^{1'}}{\prog_2} \redxs
  \config{\store^{2'}}{\varnothing}
  $$
  where $\store^2_{\vars(\prog_2)} \ne \store^{2'}_{\vars(\prog_2)}$, which is a
  contradiction.
%   This implies that there is some
%  $\store \in \mems(\vars(\prog_1';\prog_2))$ such that
%  $\store_{\vars{\prog_{1}'}} \in \runs{\prog_1'}$ and
%  \mems(\vars(\prog_1))$ with   
\end{proof}

We must also choose witnesses that allow us to generalize verification
of $\gNIMO$ preservation through gates. We need to consider separately
the definition of gates and their use. Our type discipline ensures that
gates can only be defined once, but it is also known that gate output
wire labels should only be used once \cite{XXX}, in which case we
say they are \emph{wired}. We return to this issue in more detail in
Section \ref{section-composition-copy}.
\begin{definition}
  We write $g \in \prog$ iff all views for $g$'s table are fully
  defined in $\prog$. We write $\wired(\prog)$ to denote the
  set of gates $g$ whose views occur on the rhs of an assignment
  in $\prog$.
  %$\view{1}{\ttt{gate:}g\ttt{tt}i} \in \iov(\prog)$
  %and $\view{1}{\ttt{gate:}g\ttt{pt}i} \in \iov(\prog)$ for all $0 < i \le 4$,
  %or  $\view{1}{\ttt{gate:}gi\ttt{.k}} \in \iov(\prog)$
  %and $\view{1}{\ttt{gate:}gi\ttt{.p}} \in \iov(\prog)$ for all $0 < i \le 2$.
\end{definition}
Given the above definition, the needed result follows by Lemma \ref{lemma-scope} and definition
of $\gNIMO$:
\begin{lemma}
  \label{lemma-gnimo}
  Let $X = \vars(\prog_1) \cap \vars(\prog_2)$.
  If $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
  and
  $\gNIMO(\prog_1;\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$ for
  $g \in \prog_2$,
  then $\gNIMO(\prog_1';\prog_2,\{ \ttt{k} = \be_{\ttt{k}}; \ttt{p} = \be_{\ttt{p}}\}, g)$
  for any $\prog_1'$ where $\prog_1';\prog_2$ is safe and
  $\vars(\prog_1') \cap \vars(\prog_2) = X$.
\end{lemma}

@@

Now we can demonstrate that our $\ttt{and}$ gate definition preserves
$\gNIMO$, and additionally that internale gates are probabilistically independent.
This latter fact is crucial to compositionality, and allows us for automatically
verify the circuit in isolation, with input encoding serving as an adequate input
witness for generalization. 
\begin{lemma}[Composition ($\eand$)]
  \label{lemma-compose}
Given well-typed $\prog_1;E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]$ where:
$$
\config{\prog_1}{E[\ttt{andgg}(g,g_1,g_2,v_1,v_2)]} \redxs \config{\prog_1;\prog_2}{E[v]}
$$
If the following preconditions hold:
\begin{enumerate}
\item $(\{ g_1 \} \sqcup \{ g_2 \}) \cap \wired(\prog_1) = \varnothing$
\item $\gNIMO(\prog_1,v_1,g_1) \wedge \gNIMO(\prog_1,v_2,g_2)$ 
\end{enumerate}
then we have as postconditions:
\begin{enumerate}
  \item $\gNIMO(\prog_1;\prog_2,v,g)$
  \item $(\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}) * (\margd{\progd(\prog_1;\prog_2)}{\iov(\prog_2)})$
\end{enumerate}
\end{lemma}
\begin{proof}
  We have automatically verified the postconditions (1) and (2) for the particular
  program $\prog_1;\prog_2$ resulting from execution of:
  $$e\quad \defeq\quad \ttt{let gs = encode(s1,s2) in andgg(0,s1,s2,gs.wv1,gs.wv2)}$$
  where:
  $$
  \config{\varnothing}{e} \redxs
  \config{\prog_1}{\ttt{andgg}(\ttt{0},\ttt{s1},\ttt{s2},v_1,v_2)} \redxs \config{\prog_1;\prog_2}{v}
  $$
  By Lemma \ref{lemma-encode} we have that
  $\gNIMO(\prog_1,v_1,\ttt{s1}) \wedge \gNIMO(\prog_1,v_2,\ttt{s2})$. 
  And since $\ttt{andgate}$ generates unique flips per gate, and well-typedness ensures
  uniqueness of gates, we can generalize the result to replace $\prog_1$ with arbitrary $\prog_1'$
  such that $\prog_1';\prog_2$ is safe,
  applying Lemma \ref{lemma-scope} to obtain postcondition (2) and Lemma \ref{lemma-gnimo}
  to obtain postcondition (1). 
\end{proof}
Finally, we can demonstrate that decoding correctly casts a garbled output into the
clear, and thus establishes $\NIMO$ for the garbled circuit that precedes it.
\begin{lemma}[Decoding]
  \label{lemma-decode}
Given well-typed $\prog_1;E[\ttt{decode}(g,v)]$ where
$$
\config{\prog_1}{E[\ttt{decode}(g,v)]} \redxs \config{\prog_1;\prog_2}{E[\be]}
$$
If $\gNIMO(\prog_1,v,g)$, then $\NIMO(\prog_1;\prog_2;\eassign{\outv}{\be})$.
\end{lemma}
\begin{proof}
  We have automatically verified $
  \progd(\prog,\be)(\{ \itv \mapsto 1 \}) = 1$ where:
  $$\config{\varnothing}{\ttt{decode(g,\{ k = flip[2, gate:g.k]; p = flip[2,gate:g.p] \})}}
    \redxs \config{\prog}{\be}$$
    and also $
  \progd(\prog,\be)(\{ \itv \mapsto 0 \}) = 1$ where:
  $$\config{\varnothing}{\ttt{decode(g,\{ k = not flip[2, gate:g.k]; p = not flip[2,gate:g.p] \})}}
    \redxs \config{\prog}{\be}$$
  The result follows by definition of $\gNIMO$ and $\NIMO$.
\end{proof}

\begin{fpfig}[t]{YGC copy gate definitions.}{fig-ygc-copy}
  {\footnotesize
    \begin{verbatimtab}
      sharetab2(gid, tid, k, p, b)
      {
        let r1 = k xor b in
        let r0 = (not k) xor (not b) in
        v[1,gate: || gid || tid || 1] := select(p,r1,r0);
        v[1,gate: || gid || tid || 2] := select(not p,r1,r0);
      }

      copygate(ca,cb,g)
      {
        let wl = owl(g) in
        let owl1 = owl(ca) in
        let owl2 = owl(cb) in
        sharetab2(ca,tt,wl.k,wl.p,owl1.k); sharetab2(ca,pt,wl.k,wl.p,owl1.p);
        sharetab2(cb,tt,wl.k,wl.p,owl2.k); sharetab2(cb,pt,wl.k,wl.p,owl2.p)
      }

      evalcopy(ca,cb,wv)
      {
        let wv1k = wv.k xor select(wv.p,v[1,gate: || ca || tt1], v[1,gate: || ca || tt2]) in
        let wv1p = wv.k xor select(wv.p,v[1,gate: || ca || pt1], v[1,gate: || ca || pt1]) in
        let wv2k = wv.k xor select(wv.p,v[1,gate: || cb || tt1], v[1,gate: || cb || tt2]) in
        let wv2p = wv.k xor select(wv.p,v[1,gate: || cb || pt1], v[1,gate: || cb || pt1]) in
        { wv1 = { k = wv1k; p = wv1p }; wv2 = { k = wv2k; p = wv2p } }  
      }

      copy(ca,cb,g,wv) { copygate(ca,cb,g); evalcopy(ca,cb,wv) } 
    \end{verbatimtab}
  }
\end{fpfig}

\subsection{Building and Extending Automatically Secure Circuits}
\label{section-composition-copy}

Our results in Section \ref{section-pre-post} and Theorem
\ref{theorem-nimo} imply that any well-typed program using
$\ttt{encode}$, $\ttt{andgg}$, and $\ttt{decode}$ will generate a
circuit that is passive secure, provided that gate outputs are
not used more than once. To add an $\ttt{or}$ gate to our
library, we can use $\ttt{andgg}$ as a template, in fact simple
modification of $\ttt{andtable}$ is all that would be needed. Thus,
this library embodies a Fairplay-like language where well-typedness
of metaprograms guarantees safety and security of generated protocols.

The requirements that output wire labels are not used more than once
reflects a known result, that naive copy gate output is unsound in
YGC- i.e., passive security fails. In fact, we are unable to
automatically verify $\gNIMO$ for $\ttt{andgg}$ without the assumption
that input wire lables are in independent uniform distributions. We
can capture the necessary fix as in Figure \ref{fig-ygc-copy}. The
garbler creates a copy gate with two new output wire labels in
independent uniform distributions (in \ttt{copygate}), and permutes
them using the input wire label pointer after encrypting with the
input wire label keys. The evaluator then recovers the encrypted
output wire values during evaluation (in \ttt{evalcopy}). Given an
$\ttt{orgg}$ gate defined as outlined abotve, we can then use
secure $\ttt{copy}$ as follows: 
\begin{verbatimtab}
  let s = encode(s1,s2) in
  let s1c = copy(c1s1,c2s1,s1,s.wv1) in
  let s2c = copy(c1s2,c2s2,s2,s.wv2) in
  let wv1 = andgg(1,c1s1,c1s2,s1c.wv1,s2c.wv1) in
  let wv2 = orgg(2,c2s1,c2s2,s1c.wv2,s2c.wv2) in
  decode(3,andgg(3,1,2,wv1,wv2))
\end{verbatimtab}
These definitions of $\ttt{copy}$ and $\ttt{orgg}$ have been
automatically verified as in Section \ref{section-pre-post} for
$\ttt{andgg}$. In general, any new gate form- such as optimized gates
\cite{XXX}- can be verified automatically in a similar way.

While the copy gate supports secure programming patterns, security
still relies on programmer discipline, i.e., not wiring a gate more
than once. However, we can keep track of wired gates in type effects,
similar to the manner in which we keep track of defined views- type
dependence in our system is able to accurately precisely track
gate identifiers. This is also reminiscent of the use of linearity
in other related systems for enforcing obliviousness \cite{darais2019language}.

### 2023-06-18 popl24/minicat.tex
checkpoint

We let $X$ range over sets of variables and $S,V,F$ to range over sets
of secrets, views, and flips (including invocations of the oracle)
respectively. Given a program $\prog$, we write $\iov(\prog)$ to
denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
and $V$ is the set of views in $\prog$, and we write $\flips(\prog)$
to denote the set $F$ of flip variables in $\prog$. For any set of
variables $X$ and parties $P$, we write $X_P$ to denote the subset of
$X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to denote
the subsets belonging to honest and corrupt parties, respectively.

>>>>

We let $X$ range over sets of variables and $S,V,F$ to range over sets
of secrets, views, and flips (including invocations of the oracle)
respectively. Given a program $\prog$, we write $\iov(\prog)$ to
denote $S \cup V$ where $S$ is the set of secret variables in $\prog$
and $V$ is the set of views in $\prog$, we write $\flips(\prog)$
to denote the set $F$ of flip variables in $\prog$, and we
write $\vars(\prog)$ to denote $\iov(\prog) \cup \flips(\prog)$. For any set of
variables $X$ and parties $P$, we write $X_P$ to denote the subset of
$X$ owned by any party in $P$, and we write $X_H$ and $X_C$ to denote
the subsets belonging to honest and corrupt parties, respectively.
